---
title: "Using Quarto to Create a Variety of Documents"
subtitle: "Automobile Registrations by State, 2020"
author: "Jacques Clerville"
date: "June 12, 2025"
format: html
editor: visual
---

## Load Packages

-   You have to install your packages before you load them. To install them, you can use the following code: install.packages ("**name_of_your_package**"). A package is installed once but needs to always be loaded before each use.

```{r}
#| label: load-packages
#| warning: false # Turn off warning messages
#| message: false # Turn of messages

library(tidyverse) # Contains dplyr for data manipulation, ggplot2 for data visualization, readr for csv files, and more
library(scales) # For nice labels in charts
library(ggrepel)    # For non-overlapping labels
library(ggtext)     # For fancier text handling
library(sf)         # Simple features package for importing and manipulating spatial data
library(tigris)     # To shift non-contiguous states (Alaska, Hawaii) and Puerto Rico and other functionalities
library(ggThemeAssist) # To add title, subtitle, caption to plots, and more
```

# Import Your Datasets

```{r}
#| label: import-datasets
#| warning: false # turn off warning messages
#| message: false

us_states <- read_sf("data/cb_2023_us_state_20m/cb_2023_us_state_20m.shp") # data from the US Census Bureau
Motor_vehicle <- read_csv("data/Motor_Vehicle_Registrations_Dashboard_data.csv") # We can also use read.csv function, which is included in Base R; data accessed from data.gov
```

## Create Basic US Map from US Census Bureau Data

```{r}
#| label: create-basic-US-plot
#| warning: false # turn off warning messages
#| message: false

ggplot() +
  geom_sf(data = us_states) +
  coord_sf(crs = st_crs("ESRI:102003")) +  # Albers (specific projection for the contiguous US and often used for US maps; can distort shapes for non-contiguous states)
  theme_void()
```

## Shift Alaska, Hawaii, and Puerto Rico with Shift_Geometry Function

```{r}
#| label: shift-us-map
#| warning: false
#| message: false


us_states_shifted <- us_states %>%
  shift_geometry() # function to shift non-contiguous states and PR; is included in tigris package

ggplot() +
  geom_sf(data = us_states_shifted) +
  coord_sf(crs = st_crs("ESRI:102003")) +  # Albers projection works best when non-contiguous states are shifted
  theme_void()
```

## Filter and Clean Your Datasets

-   R is **case sensitive**. Make sure you enter the value names as they appear in the dataset when you are re-coding them and use quotations for categorical values.

```{r}
#| label: filter-clean-datasets
#| warning: false
#| message: false

us_states_new <- us_states_shifted %>%
  filter (NAME != 'Puerto Rico') 

Motor_vehicle_2020 <- (Motor_vehicle) %>%
    filter(year == 2020) %>% 
    select (state, Auto, Bus, Truck, Motorcycle) %>% 
    mutate(state = case_when(
    state == "Connecticut (2)" ~ "Connecticut",
    state == "Dist. of Col." ~ "District of Columbia",
    state == "Illinois (2)" ~ "Illinois",
    state == "Kansas (2)" ~ "Kansas",
    state == "Maine (2)" ~ "Maine",
    state == "Massachusetts (2)" ~ "Massachusets",
    state == "Michigan (2)" ~ "Michigan",
    state == "Minnesota (2)" ~ "Minnesota",
    state == "Missouri (2)" ~ "Missouri",
    state == "Nevada (2)" ~ "Nevada",
    state == "New Jersey (2)" ~ "New Jersey",
    state == "Oklahoma (2)" ~ "Oklahoma",
    state == "Vermont (2)" ~ "Vermont",
    state == "Washington (2)" ~ "Washington",
    state == "West Virginia (2)" ~ "West Virginia",
    state == "Wisconsin (2)" ~ "Wisconsin",
    TRUE ~ state))  # Keep other values unchanged
```

# Merge Coordinates Dataset with Motor Vehicle Dataset (Left Join)

```{r}
#| label: merge-datasets
#| warning: false
#| message: false


combined_dataset <- us_states_new %>% 
  left_join(Motor_vehicle_2020, by = c("NAME" = "state"))
```

## Create US Map Showing Scale of Automobile Registrations by State

```{r}
#| label: create-map
#| warning: false
#| message: false

ggplot() + 
  geom_sf(data = combined_dataset, 
          aes(fill = Auto),
          linewidth = 0.25) +
  coord_sf(crs = st_crs("ESRI:102003")) +  # Albers
    scale_fill_viridis_c(option = "inferno", labels = label_comma()) +
labs(fill = "Auto")  # inferno is colorblind-friendly
```

# Save Plot

```{r}
#| label: save-map
#| warning: false
#| message: false


save_map <- ggplot() + 
  geom_sf(data = combined_dataset, 
          aes(fill = Auto),
          linewidth = 0.25) +
  coord_sf(crs = st_crs("ESRI:102003")) +  # Albers
    scale_fill_viridis_c(option = "inferno", labels = label_comma()) +
  labs(fill = "Auto")
```

# Final Plot

```{r}
#| label: add-title-subtitle-caption
#| warning: false
#| message: false


save_map + theme(plot.title = element_text(size = 16)) +labs(title = "Automobile Registrations by State, United States, 2020") + theme(plot.subtitle = element_text(size = 12)) +labs(subtitle = "California, Texas, and Florida had the highest number of registered automobiles") + theme(plot.caption = element_text(size = 10)) +labs(caption = "Source: US Census Bureau and US Government's Open Data")
```

## Resources

-   [Quarto Website](https://quarto.org)
-   [Making cleaner, nicer Quarto documents](https://datavizs24.classes.andrewheiss.com/news/2024-06-11_cleaner-nicer-qmd-output.html)
-   [US Census Bureau Data](https://www.census.gov/data.html)
-   [The Home of the U.S. Government's Open Data](https://data.gov/)
